---
title: "Exercise 03"
subtitle: "Fortgeschrittene Statistische Software für NF - SS 2022/23"
author: "Patrick Bitzer (11551540)"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


## Exercise 1: Initializing git

### Exercise 1d

strengths

* git is a version control tool, that allows to jump back in the development process and restore old versions of a file, if errors have occured. Version control also allows other people involved in a project to track changes. When developing software this is incredibally useful, as it allows other collaborating project members to examine, what changed since their last login, giving an opportunity easily get an overview over the often chaotic development process. Therefore git also allows developers to be working on the same version of the software, by ensuring that everyone is up-to-date.

* Git is designed to be incredibly efficient and fast. It uses a highly compressed and optimized data storage format, which minimizes the size of the repository on disk and speeds up operations like cloning, committing, branching, and merging. Git's architecture and algorithms are optimized to handle large repositories and scale well with the growth of a project. It is also free and open source, making it an enormous advantage for companies, as they do not need to develop their own software, nor their own server infrastructure

weaknesses

* fully mastering git-fu takes lots of efford to accomplish. Furthmore at first it is usually very difficult for noobs to get into the git basics.

* Git is primarily designed for version control of source code and text-based files. While Git can handle binary files, including large data files, it is not well-suited for managing them efficiently. Large data files can significantly increase the size of the Git repository. Git stores a complete copy of each version of every file in the repository, leading to limited scalability. Git is not designed to store repositories with large files, especially those that frequently change, as git needs to compare and iterate over both the old and the new file, every time that changes in such a big file are commited.


## Exercise 2

### Exercise 2a

https://github.com/patrickab/exeRcise-sheet-3


## Exercise 3

### Exercise 3b

```{r}
library("tidyverse")

vornamen_2021 <- read_csv(file.path("data", "vornamen_2021.csv"))
vornamen_2022 <- read_csv(file.path("data", "vornamen_2022.csv"))

typeof(vornamen_2021$Anzahl)
typeof(vornamen_2022$Anzahl)
```

the values in the "Anzahl" column are stored as character strings instead of numeric values, as the CSV datatype typically stores tabular data (numbers and text) in plain text. Therefore, all column values are by default read in as character strings.

```{r}
vornamen_2021$Anzahl <- as.integer(vornamen_2021$Anzahl)
vornamen_2022$Anzahl <- as.integer(vornamen_2022$Anzahl)

vornamen_2021 %>% head(10)
vornamen_2022 %>% head(10)
```

In both datasets the column "Anzahl" is now stored as integer datatype. Converting these string values to integer values now allows us to perform arithmetic operations on the data.

```{r}
vornamen_2021 %>% tail(10)
vornamen_2022 %>% tail(10)
```

However, since the string value "4 oder weniger" cannot be converted to a integer datatype, therefore each column-value presenting this value has been converted to "NA". In order to not loose data and keep working with at least inaccurate results, these values can be replaced with an integer of value 4. This will result in inaccurate results, as with a probability bordering on certainty not all of those NA's equal 4 in the real world. This leads to inaccuracy of results, and possibly even to incomparability of the datasets, especially if the real world average value of those NA's differs between both datasets and/or if the amount of NA's differs significantly.

```{r}
vornamen_2021$Anzahl <- ifelse(is.na(vornamen_2021$Anzahl), as.integer(4), as.integer(vornamen_2021$Anzahl))
vornamen_2022$Anzahl <- ifelse(is.na(vornamen_2022$Anzahl), as.integer(4), as.integer(vornamen_2022$Anzahl))

vornamen_2021 %>% tail(10)
vornamen_2022 %>% tail(10)
```


### Exercise 3c

The sum calculated below describes the maximum total number of babies born in 2022 and 2021 according to the dataset provided by the city of Munich. (https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen)

```{r}
total_2021 <- sum(vornamen_2021$Anzahl, na.rm = TRUE)
total_2022 <- sum(vornamen_2022$Anzahl, na.rm = TRUE)
```

Babies born in Munich, 2021: `r format(total_2021, scientific = FALSE)`

Babies born in Munich, 2022: `r format(total_2022, scientific = FALSE)`

According the dataset, and taking speciied inaccuracies into account, in Munich more babies were born in 2021, than in 2022.

However, in order to determine, which year presented the bigger baby-boom, the increase of birth relative to the previous year needs to be calculated.

```{r}
vornamen_2020 <- read_csv(file.path("data", "vornamen_2020.csv"))
vornamen_2020$Anzahl <- as.integer(vornamen_2020$Anzahl)
vornamen_2020$Anzahl <- ifelse(is.na(vornamen_2020$Anzahl), as.integer(4), as.integer(vornamen_2020$Anzahl))
total_2020 <- sum(vornamen_2022$Anzahl, na.rm = TRUE)

increase_2021 <- (total_2021/total_2020 - 1) * 100
increase_2022 <- (total_2022/total_2021 - 1) * 100
```

Increase of births in Munich, 2021 (relative to previous year): `r format(increase_2021, scientific = FALSE)`%

Increase of births in Munich, 2022 (relative to previous year): `r format(increase_2022, scientific = FALSE)`%

In 2021 total births increased, in 2022 total births decreased. Therefore 2021 presented the bigger baby-boom.


### Exercise 3d

```{r}
vornamen_2021 <- vornamen_2021 %>% mutate(year = as.integer(2021))
vornamen_2022 <- vornamen_2022 %>% mutate(year = as.integer(2022))

vornamen_2021 %>% head(10)
vornamen_2022 %>% head(10)
```

### Exercise 3e

```{r}
vornamen_combined <- bind_rows(vornamen_2021, vornamen_2022)
vornamen_combined %>% head(10)
```

### Exercise 3f

```{r}
combined_counts_top10 <- vornamen_combined %>% 
                         group_by(Vorname) %>% 
                         summarize(combined_counts = sum(Anzahl)) %>%
                         arrange(desc(combined_counts)) %>% 
                         head(10)

library("knitr")

combined_counts_top10 %>% kable(caption = "Top 10 Names of 2021 and 2021 in Munich")
```


## Exercise 4

### Exercise 4a

#### Exercise 4a Text Prompt

For this task: Specifically use ChatGPT to solve the task and submit your prompts in addition to the
solution
a) The code below does not work because the wrong apply function has been used. Find out which apply
function would be correct and why it did not work. Correct the code. Also calculate the rowwise
means.
\###Create a sample data frame
tax_data <- data.frame( Name = c(“Munich GmbH”, “ABC Inc.”, “Backpacks 1980”, “Bavarian Circus”),
Tax_2019 = c(5000, 4000, 6000, 3500), Tax_2020 = c(4800, 4200, 5800, 3700), Tax_2021 = c(5200, 3800,
5900, 3400) )
\###Calculate column-wise means
column_means <- lapply(tax_data[, -1], 2, mean)
column_means

#### Exercise 4a Answer

The code provided is using the wrong apply function. Currently, it is using lapply() instead of sapply() to calculate the column-wise means. The lapply() function applies a given function to each element of a list or vector, but it does not simplify the result. On the other hand, sapply() is a variant of lapply() that simplifies the result to a vector or matrix if possible.

To correct the code and calculate the column-wise means, you should replace lapply() with sapply().

Here's the corrected code:

```{r}
tax_data <- data.frame(
  Name = c("Munich GmbH", "ABC Inc.", "Backpacks 1980", "Bavarian Circus"),
  Tax_2019 = c(5000, 4000, 6000, 3500),
  Tax_2020 = c(4800, 4200, 5800, 3700),
  Tax_2021 = c(5200, 3800, 5900, 3400)
)

# Calculate column-wise means
column_means <- sapply(tax_data[, -1], mean)
column_means

```
Additionally, if you want to calculate the row-wise means, you can use the rowMeans() function as follows:

```{r}
# Calculate row-wise means
row_means <- rowMeans(tax_data[, -1])
row_means
```
```{r}
head(tax_data,10)
```

```{r}
row_means <- rowMeans(tax_data[, -1])
row_means
```


#### Exercise 4a Second Text Prompt

how can i calculate rowwise means and add the "Name" of each row as column of the resulting object?

#### Exercise 4a Second Answer

```{r}
row_means <- rowMeans(tax_data[, -1])
row_means_with_name <- cbind(Name = tax_data$Name, Mean = row_means)

row_means_with_name
```

### Exercise 4b

#### Exercise 4b Text Prompt

in RStudio, explain what the rapply() function does. Create an easy example with mock data where the function is used.

#### Exercise 4b Answer

In RStudio, the rapply() function is used to apply a specified function recursively to elements of a list or nested list structure. It is particularly useful when you have a complex nested list and want to apply a function to each element in a consistent manner.

The syntax for rapply() is as follows:

rapply(object, f, classes = "ANY", how = "replace", ...)

* object represents the list or nested list structure to which the function will be applied.

* f is the function that you want to apply to each element.

* classes is an optional argument that specifies the classes to which the function will be applied. By default, it is set to "ANY" which means that the function will be applied to all elements regardless of their class.

* how determines the structure of the result. It can take three possible values: "replace" (default), "unlist", or "list".

* ... represents additional arguments that can be passed to the function f.

Now, let's create an example with mock data to illustrate the usage of `rapply()`. Suppose we have a nested list representing a hierarchical directory structure:

```{r}
directory <- list(
  folder1 = list(
    subfolder1 = list(
      file1 = "data1.csv",
      file2 = "data2.csv"
    ),
    subfolder2 = list(
      file3 = "data3.csv",
      file4 = "data4.csv"
    )
  ),
  folder2 = list(
    subfolder3 = list(
      subsubfolder1 = list(
        file5 = "data5.csv",
        file6 = "data6.csv"
      ),
      subsubfolder2 = list(
        file7 = "data7.csv",
        file8 = "data8.csv"
      )
    ),
    subfolder4 = list(
      file9 = "data9.csv",
      file10 = "data10.csv"
    )
  )
)
```

In this example, we have four folders (folder1, folder2, subfolder1, subfolder2) and two subfolders (subsubfolder1, subsubfolder2). Each folder and subfolder contains files (file1 to file10) with different names.

You can use the rapply() function to apply a function to each file in this nested directory structure For instance, let's define the function get_filename(). The get_filename() function uses the file_path_sans_ext() function from the tools package to remove the file extension. We pass this function as an argument to rapply() along with the directory object.

```{r}
get_filename <- function(file) {
  tools::file_path_sans_ext(file)
}

result <- rapply(directory, get_filename, classes = "character")

print(result)

```

#### Exercise 4b Own Words

Rapply is a function, that can be used on recursive data structures, like linked lists, binary trees, directory structures, family trees, or any other kind of recursive tree structure. `rapply(recursive_structure, function)` recursively applies a user defined `function` to each element within a given `recursive_structure`. In a binary tree for example, rapply() could be used to convert binary sorted temperature data from celsius to fahrenheit. By specifying a function accordingly (regular expressions for `[df01,...,dfn]` and `[data01.csv,...,datan.csv]`) , it could also be used to recursively load all `.csv` files from a folderstructure as dataframes into an RStudio environment, instead of manually calling

`data1337 <- read_csv(file.path("data/path/to/subfolder", "data1337.csv"))`

for every element inside the folder.